---
title: "Data Tidying"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To add a code block: cmd + option + i (for mac)
### Insert code chunk for calling packages
To get a function from a package - package::function, ex: dplyr::filter
To install a package (so that you can bring it from library) - install.packages("dplyr")

```{r, message = FALSE, warning = FALSE}
library(dplyr)
library(tidyr)
library(readr)
```

### Read in a dataset from an open source url
This is beneficial for collaboration bc other people don't have to have the dataset downloaded
```{r}
catch_original <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")
```
read_csv is part of readr and will read in the 'whole' dataset - more complete than read.csv

The pipe (%>%) operator: the output of one argument is the input of the next, think 'and then'. The pipe command has a keyboard shortcut: cmd + shift + m

## Clean Data
```{r}
catch_data <- catch_original %>% 
  select(Region, Year, Chinook, Coho, Pink, Sockeye, Chum)
  #select(-ALL, -notesRegCode)
  
head(catch_data)
```
### Change chinook from character to numeric type

```{r}
catch_clean <- catch_data %>%
  mutate(Chinook = if_else(Chinook == "I", "1", Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook))
head(catch_clean)
```
### Finding spurious characters in Chinook column
random variable i is assigned to any value that is TRUE for NA. Show the row that contains this problem character
which returns the indeces where said value (na of is.na) is TRUE
```{r}
i <- which(is.na(catch_clean$Chinook))
catch_original[i, ]
```
## Reshaping Data
wide format to long format
```{r}
catch_long <- catch_clean %>% 
  pivot_longer(cols = -c(Region, Year), names_to = "Species", values_to = "Catch") %>% 
  rename(catch_thousands = Catch) %>% #rename function
  mutate(catch = catch_thousands * 1000) %>% #add a column
  select(-catch_thousands) #deletes new column

head(catch_long)
```

```{r}
catch_wide <- catch_long %>% 
  pivot_wider(names_from = Region, values_from = catch)

head(catch_wide)
```

## Split-Apply-Combine
Calculate mean catch by region and create new table
group_by stores some information (in an attribute) that remains pretty invisible unless you go find it
select grabs columns
filter grabs rows

```{r}
mean_region <- catch_long %>% 
  filter(Species == "Sockeye") %>% 
  group_by(Region, Species) %>% 
  summarise(catch_mean = mean(catch),
            n = n(),
            .groups = "drop") #ungroup

head(mean_region)
```

## Joins
Here we want to Leftjoin, or merge the right table into the left one (catch_long)
```{r}
region_defs <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1") %>% 
    select(code, mgmtArea)

head(region_defs)
```

```{r}
catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))

head(catch_joined)
nrow(catch_joined) == nrow(catch_long) # check to make sure you joined it correctly
```

