---
title: "Rmarkdown_lesson"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Header 1

### Header 3

Write some text here. **This is in bold**

-   this makes a list
-   so does this
-   to make a list, you have to have a blank line on top of the list
-   or put two spaces after the line above it

*For any help with markdown commands, see markdown quick reference under tools tab*
Citations are also easy to insert, you can also auto-generate a bibliography at the end

## Header 2

1.  numbered list
2.  numbered list
3.  can also
4.  make an
5.  ordered list
6.  like this

so even if you forget what number you're on, you can always put a 1.

```{r}
# hitting the C+ button up to will create this automatically
# cmd + option + i is the keyboard shortcut
# or you can type it manually

x <- 3 + 2
```

Running code: several buttons to use, or control + enter on the keyboard
play button runs everything in a chunk of code, control+enter runs one line

## Read in Data

```{r}
bg_chem <- read.csv("../data/BGchem2008data.csv")
```


## Analysis


## Plots
echo=FALSE hides all code in the block in the html

```{r,echo=FALSE}
plot(bg_chem$CTD_Depth, bg_chem$CTD_Temperature)
```

